---
title: "SSI Research Project"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Setup

```{r}
library(haven)
library(readr)
library(car)
setwd("/Users/Carlos_Leo/ssiresearch")
data <- read_dta("NLSFW1W2 Dataset.dta")
```


Independent variables:
older siblings = data$osiblyr
younger siblings = data$ysiblyr

Mediating variables:
Last Year, How Often Did Your Parents Or Other Adults In Your Household Check If You’d Done Your Homework? = data$w1q38a
Meet personally with your teachers? = data$w1q38b
Help you with your homework? = data$w1q38c
Reward you for good grades? = data$w1q38e
Punish you for bad grades? = data$w1q38f

Control variables:
Did You Attend All Or Most Of Your Senior Year In = data$w1q38
Sex of Respondent = data$sex
Respondent’s Ethnicity = data$w1qzeth
Parents in household (respondent last year before college) = data$parntlyr
Overall Quality Of The School = data$w1q73k
What Is The Highest Level Of Schooling Achieved By Your Mother Or The Woman Most Responsible For Raising You? = data$w1q151
What Is The Highest Level Of Schooling Achieved By Your Father Or The Man Most Responsible For Raising You? = data$w1q152
Household income = data$w1q179
Record High School State = data$w1q53d
```{r}

for (i in 6:8) {
  data$w1q38a[which(data$w1q38a == i)] <- NA
  data$w1q38b[which(data$w1q38b == i)] <- NA
  data$w1q38c[which(data$w1q38c == i)] <- NA
  data$w1q38e[which(data$w1q38e == i)] <- NA
  data$w1q38f[which(data$w1q38f == i)] <- NA
  
  data$w1q73k[which(data$w1q73k == i)] <- NA
  
  data$w1q63a[which(data$w1q63a == i)] <- NA
  data$w1q63b[which(data$w1q63b == i)] <- NA
  data$w1q63c[which(data$w1q63c == i)] <- NA
  data$w1q63d[which(data$w1q63d == i)] <- NA
  data$w1q63e[which(data$w1q63e == i)] <- NA
  data$w1q63f[which(data$w1q63f == i)] <- NA
}

# dependent variables
for (j in 95:98) {
  data$w1q63a[which(data$w1q63a == j)] <- NA
  data$w1q63b[which(data$w1q63b == j)] <- NA
  data$w1q63c[which(data$w1q63c == j)] <- NA
  data$w1q63d[which(data$w1q63d == j)] <- NA
  data$w1q63e[which(data$w1q63e == j)] <- NA
  data$w1q63f[which(data$w1q63f == j)] <- NA
  
  data$w1q151[which(data$w1q151 == j)] <- NA
  data$w1q152[which(data$w1q152 == j)] <- NA
  data$w1q179[which(data$w1q179 == j)] <- NA
  data$w1q38[which(data$w1q38 == j)] <- NA
}

for (k in 996:998) {
  data$w1q153o[which(data$w1q153o == k)] <- NA
  data$w1q157o[which(data$w1q157o == j)] <- NA
}

data$w1q63c[which(data$w1q63c == 5)] <- NA
data$w1q63d[which(data$w1q63d == 5)] <- NA
data$w1q63f[which(data$w1q63f == 5)] <- NA

data$w1q38[which(data$w1q38 == 8)] <- 5

data$eng <- recode(data$w1q63a, "1=4; 2=3; 3=2; 4=1")

data$hist <- recode(data$w1q63b, "1=4; 2=3; 3=2; 4=1")

data$math <- recode(data$w1q63c, "1=4; 2=3; 3=2; 4=1")

data$sci <- recode(data$w1q63d, "1=4; 2=3; 3=2; 4=1")

data$socs <- recode(data$w1q63e, "1=4; 2=3; 3=2; 4=1")

data$lang <- recode(data$w1q63f, "1=4; 2=3; 3=2; 4=1")

data$numsib <- (data$osiblyr + data$ysiblyr)
data$grade <- (data$eng + data$hist + data$math + data$sci + data$socs + data$lang) / 6
data$atten <- (data$w1q38a + data$w1q38b + data$w1q38c + data$w1q38e + data$w1q38f) / 5
data$edulvl <- (data$w1q151 + data$w1q152) / 2

names(data)[names(data) == "w1q38"] <- "schooltype"
names(data)[names(data) == "w1qzeth"] <- "race"
names(data)[names(data) == "parntlyr"] <- "numparents"
names(data)[names(data) == "w1q73k"] <- "schoolqual"
names(data)[names(data) == "w1q179"] <- "income"

data$mother <- ifelse((data$numparents == 1) || (data$numparents == 3), 1, 0)
data$father <- ifelse((data$numparents == 2) || (data$numparents == 3), 1, 0)

data$gender <- ifelse(data$sex == "F", 1, 0)
data$race <- as.numeric(factor(data$race, levels = unique(data$race)))

data$schooltype <- as.numeric(data$schooltype)
```

```{r, results = 'asis'}
mod = lm(grade ~ numsib + edulvl + numparents + schooltype + schoolqual + income + race + gender, data = data)
mod1 = lm(atten ~ numsib + edulvl + numparents + schooltype + schoolqual + income + race + gender, data = data)
mod2 = lm(grade ~ numsib + atten + edulvl + numparents + schooltype + schoolqual + income + race + gender, data = data)
summary(mod)
summary(mod1)
summary(mod2)
mod3 = lm(grade ~ osiblyr + ysiblyr + atten + edulvl + numparents + schooltype + schoolqual + income + race + gender, data = data)
summary(mod3)

library(stargazer)
options(stargazer.comment = FALSE)
stargazer(mod1, mod, mod2,
          title = "Factors Affecting High School Academic Performance of Incoming College Students",
          covariate.labels = c("Number of Siblings", "Parental Attention", "Parents' Education Level",
                               "Parents in the Household", "School Type", "School Quality",
                               "Household Income", "Race", "Female", "Constant"), 
          dep.var.labels.include = F,
          column.labels = c("IV vs MV","IV vs DV", "IV + MV vs DV"),
          model.numbers = FALSE, no.space = TRUE, 
          header = FALSE, type = "latex", font.size = "small")

stargazer(mod3,
          title = "Factors Affecting High School Academic Performance of Incoming College Students",
          covariate.labels = c("Number of Older Siblings", "Number of Younger Siblings", "Parental Attention", "Parents' Education Level",
                               "Parents in the Household", "School Type", "School Quality",
                               "Household Income", "Race", "Female", "Constant"), 
          dep.var.labels.include = F,
          model.numbers = FALSE, no.space = TRUE, 
          header = FALSE, type = "latex", font.size = "small")
```

``{r}
tab <- table(data$osiblyr)
mean(data$osiblyr, na.rm = TRUE)
barplot(tab,
        main = "“Number of older brothers and sisters”",
        names.arg=c("0", "1", "2", "3", "4", "5"),
        ylab = "Number of Students",
        xlab = "Number of Siblings")
tab <- table(data$ysiblyr)
mean(data$ysiblyr, na.rm = TRUE)
barplot(tab,
        main = "“Number of younger brothers and sisters”",
        names.arg=c("0", "1", "2", "3", "4", "5", "6", "7"),
        ylab = "Number of Students",
        xlab = "Number of Siblings")

tab <- table(data$numsib)
barplot(tab,
        main = "Number of siblings",
        names.arg=c("0", "1", "2", "3", "4", "5", "6", "7", "8"),
        ylab = "Number of Students",
        xlab = "Number of Siblings")
avgnum <- mean(data$numsib, na.rm = TRUE)
avgnum


tab <- table(data$w1q38a)
barplot(tab,
        main = "“Check if you’d done your homework?”",
        names.arg=c("Never", "Rarely", "Sometimes", "Often", "Very Often"),
        ylab = "Number of Students",
        xlab = "Frequency")

tab <- table(data$w1q38b)
barplot(tab,
        main = "“Meet personally with your teachers?”",
        names.arg=c("Never", "Rarely", "Sometimes", "Often", "Very Often"),
        ylab = "Number of Students",
        xlab = "Frequency")

tab <- table(data$w1q38c)
barplot(tab,
        main = "“Help you with your homework?”",
        names.arg=c("Never", "Rarely", "Sometimes", "Often", "Very Often"),
        ylab = "Number of Students",
        xlab = "Frequency")

tab <- table(data$w1q38e)
barplot(tab,
        main = "“Reward you for good grades?”",
        names.arg=c("Never", "Rarely", "Sometimes", "Often", "Very Often"),
        ylab = "Number of Students",
        xlab = "Frequency")

tab <- table(data$w1q38f)
barplot(tab,
        main = "“Punish you for bad grades?”",
        names.arg=c("Never", "Rarely", "Sometimes", "Often", "Very Often"),
        ylab = "Number of Students",
        xlab = "Frequency")

hist(data$atten,
     main = "Overall Attention Index",
     xlab = "Number")
avgatten <- mean(data$atten, na.rm = T)
avgatten

tab <- table(data$eng)
barplot(tab,
        main = "English",
        names.arg=c("Mostly D's", "Mostly C's", "Mostly B's", "Mostly A's"),
        ylab = "Number of Students",
        xlab = "Grade Level")

tab <- table(data$hist)
barplot(tab,
        main = "History",
        names.arg=c("Mostly D's", "Mostly C's", "Mostly B's", "Mostly A's"),
        ylab = "Number of Students",
        xlab = "Grade Level")

tab <- table(data$math)
barplot(tab,
        main = "Math",
        names.arg=c("Mostly D's", "Mostly C's", "Mostly B's", "Mostly A's"),
        ylab = "Number of Students",
        xlab = "Grade Level")

tab <- table(data$sci)
barplot(tab,
        main = "Science",
        names.arg=c("Mostly D's", "Mostly C's", "Mostly B's", "Mostly A's"),
        ylab = "Number of Students",
        xlab = "Grade Level")

tab <- table(data$socs)
barplot(tab,
        main = "Social Science",
        names.arg=c("Mostly D's", "Mostly C's", "Mostly B's", "Mostly A's"),
        ylab = "Number of Students",
        xlab = "Grade Level")

tab <- table(data$lang)
barplot(tab,
        main = "Foreign Languages",
        names.arg=c("Mostly D's", "Mostly C's", "Mostly B's", "Mostly A's"),
        ylab = "Number of Students",
        xlab = "Grade Level")

hist(data$grade,
     main = "Overall Grade",
     xlab = "Number")
avggrade <- mean(data$grade, na.rm = T)
avggrade

plot(data$atten, data$grade)
``

``{r}
library(car)
crPlots(mod2, "numsib", xlab = "Number of Siblings", ylab = "Component Residuals")
crPlots(mod2, "atten", xlab = "Parental Attention", ylab = "Component Residuals")
spreadLevelPlot(mod2, main = "Studentized Residuals versus Fitted Values in the IV + MV vs DV Model")
mean(data$income, na.rm = T)
plot(data$numsib, data$w1q38b)
``


``{r}
library(ggplot2)
library(reshape2)
library(plyr)
x <- c(0, 1, 2, 3, 4, 5, 6, 7, 8)
old <- c(3080, 726, 191, 13, 2, 1, 0, 0, 0)
young <- c(1815, 1412, 517, 134, 33, 4, 6, 2, 0)
all <- c(1289, 1643, 706, 211, 51, 8, 10, 1, 4)
df <- rename(df, c("old" = "Number of older siblings", "young" = "Number of younger siblings", "all"="Total number of siblings"))
df1 <- melt(df, id.vars = 'x')
ggplot(df1, aes(x = x, y = value, fill = variable)) + geom_bar(stat = 'identity', position = 'dodge') + xlab("Number of Siblings") + ylab("Number of Students")

``


``{r}
x <- c("Never", "Rarely", "Sometimes", "Often", "Very Often")
a <- c(1428, 1240, 710, 350, 194)
b <- c(1000, 1389, 1025, 342, 133)
c <- c(1580, 1312, 768, 188, 73)
d <- c(840, 857, 1025, 706, 458)
e <- c(2257, 888, 369, 220, 146)
df <- data.frame(x, a, b, c, d, e)
df <- rename(df, c("a"="“Check if you’d done your homework?”", "b"="“Meet personally with your teachers?”", "c"="“Help you with your homework?”", "d"="“Reward you for good grades?”", "e"="“Punish you for bad grades?”"))
df1 <- melt(df, id.vars = 'x')
ggplot(df1, aes(x = x, y = value, fill = variable)) + geom_bar(stat = 'identity', position = 'dodge') + xlab("Frequency") + ylab("Number of Students")

hist(data$atten,
     main = "Overall Attention Index",
     xlab = "Number")
``

``{r}
x <- c("Mostly D's", "Mostly C's", "Mostly B's", "Mostly A's")
English <- c(1, 47, 912, 2959)
History <- c(4, 74, 910, 2926)
Math <- c(13, 193, 1212, 2498)
Science <- c(7, 104, 1054, 2713)
soc <- c(4, 50, 751, 2984)
lan <- c(10, 111, 831, 2951)
df <- data.frame(x, English, History, Math, Science, soc, lan)
df <- rename(df, c("soc"="Social Science", "lan"="Foreign Languages"))
df1 <- melt(df, id.vars = 'x')
ggplot(df1, aes(x = x, y = value, fill = variable)) + geom_bar(stat = 'identity', position = 'dodge') + xlab("Grade Level") + ylab("Number of Students")
``

